Integer Programming Formulations for the Shared Multicast Tree Problem

### 2018-09-06 smt/sec1-intro.tex
MI: (SMT) Added a reference (Guo and Yang), newly written text is now in a blue font. Old text that is moved to different sections/paragraphs remains black

@@

@@

The remainder of this paper is organized as follows: 
Section \ref{sec:literature} summarizes relevant literature and previous works.
%Section \ref{sec:SBT} describes the SMT problem and gives detailed explanation of its objective function.
Section \ref{sec:SBT} divided into three subsections introduces the notation, explains the network model, and gives formal definition of the problem.
Integer linear programming formulations, valid inequalities and their analysis are presented in Section \ref{sec:ILP},
followed by Section \ref{sec:comp}, where the models under study are compared with respect to strength.
In Section \ref{sec:strong}, the models are further extended by variables in higher dimensional spaces,
and a constraint generation procedure is developed.
Computational results are reported in Section \ref{sec:exp},
and suggestions to future work are summarized in Section \ref{sec:conclusion}.

>>>>

\textcolor{blue}{\subsection{Related problems}}

@@

The SBT and SMT problems resemble the more frequently studied \textsc{Range Assignment Problem} (RAP) in the sense that they all ask for an undirected source-independent tree of minimum power.
However, RAP concerns cases where a connected topology must be established by means of \emph{bidirectional} links,
with the feature that power must be assigned to both the sending and the receiving device in each link.
This scenario induces a spanning tree problem, where costs are incurred only for edges that are the most expensive ones incident to any of their end nodes.

\textcolor{blue}{\subsection{Contributions}}

@@

For the same reasons that MEB has been generalized to MEM, it is natural and desirable to extend SBT to its multicast equivalent.
The \textsc{Minimum Shared Multicast Tree} (SMT) problem is thus at the forefront of this paper,
and neither MEB, MEM, nor RAP will be considered further.
Focus is henceforth on instances in which some of the nodes neither initiate new messages nor need messages from others.
Utilizing such \emph{non-destination} nodes to forward messages can in many cases contribute to substantial reductions in the total power consumption.
Non-destination nodes thus have the potential to play the role of Steiner nodes in the multicast tree.
All devices that can initiate a transmission, referred to as \emph{destination} nodes, also have to receive every message.

\textcolor{blue}{\subsection{Structure of the paper}}

The remainder of this paper is organized as follows: 
\textcolor{blue}{Section \ref{sec:literature} summarizes relevant literature and previous works.}
%Section \ref{sec:SBT} describes the SMT problem and gives detailed explanation of its objective function.
\textcolor{blue}{Section \ref{sec:SBT} divided into three subsections introduces the notation, explains the network model, and gives formal definition of the problem.}
Integer linear programming formulations, valid inequalities and their analysis are presented in Section \ref{sec:ILP},
followed by Section \ref{sec:comp}, where the models under study are compared with respect to strength.
In Section \ref{sec:strong}, the models are further extended by variables in higher dimensional spaces,
and a constraint generation procedure is developed.
Computational results are reported in Section \ref{sec:exp},
and suggestions to future work are summarized in Section \ref{sec:conclusion}.

### 2018-09-06 smt/sec2-literature.tex
MI: (SMT) Added a reference (Guo and Yang), newly written text is now in a blue font. Old text that is moved to different sections/paragraphs remains black

\section{Literature Overview}
\label{sec:literature}
The combinatorial optimization literature is already rich on Integer Programming (IP) approaches to power minimization in wireless networks.
Since the work of \citet{Wieseltier00onthe}, in particular the MEB and MEM problems have been studied extensively.
Noteworthy contributions to this research include the IP models studied by \citet{das03},
\citet{altinkemer05}, \citet{leggieri08}, \citet{yuan05}, \citet{yuan08}, \citet{bauer08}, and \citet{montemanni11}.
\citet{cagalj02} proved that MEB, and thereby MEM, is NP-hard.
Numerous contributions to fast, but possible inexact methods for MEM have thus appeared, and the interested reader is referred to \citep{hsiao13} for an overview.

@@

The SBT and SMT problems resemble the more frequently studied \textsc{Range Assignment Problem} (RAP) in the sense that they all ask for an undirected source-independent tree of minimum power.
However, RAP concerns cases where a connected topology must be established by means of \emph{bidirectional} links,
with the feature that power must be assigned to both the sending and the receiving device in each link.
This scenario induces a spanning tree problem, where costs are incurred only for edges that are the most expensive ones incident to any of their end nodes.
Simple instances \citep{Haugland12Dual} show that because of the difference in objective functions, the optimal solutions to RAP and SBT can differ substantially.
\citet{kirousis97} and \citet{clementi99} conduct
hardness results for instances of RAP embedded in three- and two-dimensional Euclidean space, respectively.
Later, IP models have been developed by e.g.\ \citet{althaus03}, \citet{montemanni04}, \citet{das05}, and \citet{Haugland11Compact}.

For the same reasons that MEB has been generalized to MEM, it is natural and desirable to extend SBT to its multicast equivalent.
The \textsc{Minimum Shared Multicast Tree} (SMT) problem is thus at the forefront of this paper,
and neither MEB, MEM, nor RAP will be considered further.
Focus is henceforth on instances in which some of the nodes neither initiate new messages nor need messages from others.
Utilizing such \emph{non-destination} nodes to forward messages can in many cases contribute to substantial reductions in the total power consumption.
Non-destination nodes thus have the potential to play the role of Steiner nodes in the multicast tree.
All devices that can initiate a transmission, referred to as \emph{destination} nodes, also have to receive every message.

>>>>

\section{Literature Overview}
\label{sec:literature}
The combinatorial optimization literature is already rich on Integer Programming (IP) approaches to power minimization in wireless networks.
Since the work of \citet{Wieseltier00onthe}, in particular the MEB and MEM problems have been studied extensively.
Noteworthy contributions to this research include the IP models studied by \citet{das03},
\citet{altinkemer05}, \citet{yuan05}, \citet{yuan08}, \citet{bauer08}, \citet{montemanni11} and \citet{leggieri08}. 
\textcolor{blue}{
The latter paper proposes an IP formulation of MEM based on a Set Covering model together with a preprocessing procedure reducing the number of constraints.}
\citet{cagalj02} proved that MEB, and thereby MEM, is NP-hard.
Numerous contributions to fast, but possible inexact methods for MEM have thus appeared, and the interested reader is referred to \citep{hsiao13} for an overview. 
\textcolor{blue}{Another study concerning MEM is \cite{guo}, where the authors investigate energy conservation while using adaptive antennas. 
Unlike omnidirectional antennas, adaptive antennas can concentrate their transmission power in a smaller angle.}

@@

Simple instances \citep{Haugland12Dual} show that because of the difference in objective functions, the optimal solutions to RAP and SBT can differ substantially.
\citet{kirousis97} and \citet{clementi99} conduct
hardness results for instances of RAP embedded in three- and two-dimensional Euclidean space, respectively.
Later, IP models have been developed by e.g.\ \citet{althaus03}, \citet{montemanni04}, \citet{das05}, and \citet{Haugland11Compact}.

### 2018-09-06 smt/sec2-sbtsmt.tex
MI: (SMT) Suggestions from Reviewer 2 - Abstract shortened from 22 to 15 lines. Introduced Section 2 Literature overview, although nothing has been changed yet in this section. Section 3 divided into 3 subsections - Assumptions and notation, Network model, Problem definition. The last paragraph in Section 1 (Introduction) is updated according to the new structure of Section 2 and 3.
MI: (SMT) Added a reference (Guo and Yang), newly written text is now in a blue font. Old text that is moved to different sections/paragraphs remains black

@@

@@

Different energy models can be found in the literature. 
\cite{halgamuge} present a comprehensive energy model for wireless sensor networks, 
where the sources of energy consumption consist of sensing, logging, processing, transimission, actuation, transient and cluster formation.
Our energy model focuses on the energy consumed by transmission, which is also tho most energy consuming aspect of wireless networks (51\% according to \cite{halgamuge}).
Some of the measures such as sensing and cluster formation are not relevant in our model.
Another distinction is that nodes in our model are all equivalent and the data is not aggregated in the nodes.
For that reason we disregard the potential energy consumption sources occuring in sensor networks.
However, it is easy to incorporate some of the disregarded measures by adding a constant for every used Steiner node.

>>>>

\textcolor{blue}{\subsection{Network model} \label{sec:netmod}}

@@

\textcolor{blue}{
Different energy models can be found in the literature. 
\cite{halgamuge} present a comprehensive energy model for wireless sensor networks, 
where the sources of energy consumption consist of sensing, logging, processing, transimission, actuation, transient and cluster formation.
Our energy model focuses on the energy consumed by transmission, which is also tho most energy consuming aspect of wireless networks (51\% according to \cite{halgamuge}).
Some of the measures such as sensing and cluster formation are not relevant in our model.
Another distinction is that nodes in our model are all equivalent and the data is not aggregated in the nodes.
On top of that, all the related works menioned in Section \ref{sec:literature} assume transmission as the sole energy consumption source. 
For these reasons we disregard the potential energy consumption sources occuring in sensor networks.}

\textcolor{blue}{
However, it is easy to incorporate some of the disregarded measures by adding a constant for every used Steiner node.}

@@
